Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2017-02-13 overall.tex
updates to the I/O section per Sergey's feedback and new graphs

The ATLAS framework has a configuration and initialization stage, which is
called Athena. At this stage, the running job is assembled  on the fly from a
large number of shared libraries.  Also, at this stage, every algorithm and
service is being configured, at run time, by a corresponding set of Python
scripts, which results in a large number of read I/O accesses to small python
scripts, with many includes and imports Python calls.

When ATLAS was originally deployed at the OLCF, all shared libraries were
placed at the Spider 2 Lustre file system. However, it was quickly discovered
that the I/O pattern at the Athena stage was detrimental to the overall Spider
2 file system performance. Since Spider 2 is a center-wide file system and
shared by all OLCF resources and users, this resulted in lower overall
interactive and metadata performance at OLCF\@. As can be seen in
Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
corresponding to the beginning of the runs (i.e., the Athena stage) before
tapering off.

@@

Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
improvement on Titan. The circled region illustrates the switch from Lustre to
NFS-exported directory for hosting the ATLAS release.

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda-completed-jobs-sw-move.pdf}}\\
    \end{tabular}
    \caption{ATLAS performance improvement on Titan. The circled region shows the switch from Lustre to NFS-exported directory for hosting the ATLAS release.}
\label{fig:atlas-perf-improvement}
\end{figure}

@@

ATLAS jobs are file read I/O heavy, as can be seen in
Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
aggregate in this observed period was less than 250 GB and the maximum amount
written in aggregate was less than 75 GB\@. The average amount read per job is
20 GB and average amount written is 6 GB\@.

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file read I/O histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-io-read}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-io-written}
\end{figure}

Per job read and write file I/O statistics show an interesting pattern.
Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
compute node on Titan is less than 400 MB on average, while the amount of data
written per node is less than 170 MB on average. This correlates with our
finding that ATLAS PanDA jobs are read heavy. However, as can be seen in
Table~\ref{panda-olcf-stats} and figures {fig:atlas-titan-io-read} and
{fig:atlas-titan-io-written}, the distribution between read and written amount
of data per job is quite different from one another. The read I/O distribution
per job shows a long tail, ranging from 25 GB to 250 GB, while the written
amount of data has a very narrow distribution.

@@

On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file $open()$
operations and 5 file $close()$ operations per second. As can be seen from
figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close},
they exhibit similar distributions. The maximum number of file $open()$
operations are around 170 on average and the maximum number of file $close()$
operations are 39 on average per second per job. The total number of file
$open()$ operations is 172,089,760 for the observed window of 174 hours, while
the total number of file $close()$ operations stand at 40,132,992 for the total
1,175 ATLAS PanDA jobs in the same observation window. The difference between
these two values is puzzling and it is under investigation at the time of
writing this paper. {\sarp{can someone take a look at this? I would expect at
least similar values for number of opens and closes and there is vast
difference. Can it be that ATLAS jobs do a lot of opens without following
closes, or perhaps multiple opens for the same files over and over?}}

>>>>

Athena, the ATLAS framework has a configuration and initialization stage. At
this stage, the running job is assembled  on the fly from a large number of
shared libraries.  Also, at this stage, every algorithm and service is being
configured, at run time, by a corresponding set of Python scripts, which
results in a large number of read operations accesses to small python scripts, with
many includes and imports Python calls.

When ATLAS was originally deployed at the OLCF, all shared libraries were
placed at the Spider 2 Lustre file system. However, it was quickly discovered
that the I/O pattern at the Athena stage was detrimental to the overall Spider
2 file system performance. Since Spider 2 is a center-wide file system and
shared by all OLCF resources and users, this resulted in lower overall
interactive and metadata performance at OLCF\@. As can be seen in
Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
corresponding to the beginning of the runs before tapering off.

@@

%Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
%improvement on Titan. The circled region illustrates the switch from Lustre to
%NFS-exported directory for hosting the ATLAS release.

%\begin{figure}[!htb]
%    \centering
%    \begin{tabular}{cc}
%        {\includegraphics[width=0.48\textwidth]{figures/panda-completed-jobs-sw-move.pdf}}\\
%    \end{tabular}
%    \caption{ATLAS performance improvement on Titan. The circled region shows the switch from Lustre to NFS-exported directory for hosting the ATLAS release.}
%\label{fig:atlas-perf-improvement}
%\end{figure}

@@

ATLAS jobs issues a large number of file read operations, as can be seen in
Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
aggregate in this observed period was less than 250 GB and the maximum amount
written in aggregate was less than 75 GB\@. The average amount read per job is
20 GB and average amount written is 6 GB\@.

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-io-read}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-io-written}
\end{figure}

Per job read and write file I/O statistics show an interesting pattern.
Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
compute node on Titan is less than 400 MB on average, while the amount of data
written per node is less than 170 MB on average. This correlates with our
finding that ATLAS PanDA jobs are read heavy. However, as can be seen in
Table~\ref{panda-olcf-stats} and figures {fig:atlas-titan-io-read} and
{fig:atlas-titan-io-written}, the distribution between read and written amount
of data per job is quite different from one another. The read operation distribution
per job shows a long tail, ranging from 12.5 GB to 250 GB, while the written
amount of data has a very narrow distribution.

@@

On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file $open()$
operations and 5 file $close()$ operations per second. The file $open()$
operations listed here don't include the file $stat()$ operations. As can be
seen from figures~\ref{fig:atlas-titan-file-open}
and~\ref{fig:atlas-titan-file-close}, they exhibit similar distributions. The
maximum number of file $open()$ operations are around 170 on average and the
maximum number of file $close()$ operations are 39 on average per second per
job. The total number of file $open()$ operations is 172,089,760 for the
observed window of 174 hours, while the total number of file $close()$
operations stand at 40,132,992 for the total 1,175 ATLAS PanDA jobs in the same
observation window. The difference between these two values is puzzling and it
is under investigation at the time of writing this paper. One possible
explanation is that ATLAS PanDA jobs perhaps don't call a file $close()$
operation per every file $open()$ issued.
